### 2021-07-09 misc/collapse cheat sheet/collapse cheat sheet.tex

\documentclass[10pt,a4paper,landscape]{article}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{multicol,multirow}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[landscape]{geometry}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{booktabs}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[scaled=1]{helvet}
\renewcommand\familydefault{\sfdefault} 
\usepackage[T1]{fontenc}

\usepackage{setspace} % \linespread{1.3}
\setlength\parindent{0pt}

\usepackage{graphicx}
\usepackage[section]{placeins}
\usepackage{adjustbox}

% \title{\textcolor{Gray}{\textbf{POLICY BREIF}}\\ \vspace{2mm} \textcolor{RoyalBlue}

\ifthenelse{\lengthtest { \paperwidth = 11in}}
    { \geometry{top=.2in,left=.3in,right=.3in,bottom=.3in} } %Normally all .5in (L,R=4 before)
	{\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
		{\geometry{top=0.8cm,left=1cm,right=1cm,bottom=1cm} }
		{\geometry{top=1cm,left=1cm,right=1cm,bottom=1cm} }
	}
\pagestyle{empty}
\makeatletter
\renewcommand{\section}{\@startsection{section}{1}{0mm}%
                                {-2ex plus -.5ex minus -.2ex}%before -1
                                {0.5ex plus .2ex}%x
                                {\normalfont\large\bfseries}}
\renewcommand{\subsection}{\@startsection{subsection}{2}{0mm}%
                                {-2explus -.5ex minus -.2ex}%before -1
                                {0.5ex plus .2ex}%
                                {\normalfont\normalsize\bfseries}}
\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{0mm}%
                                {-2ex plus -.5ex minus -.2ex}%set spaces around sections
                                {0.3ex plus .2ex}%normally 1
                                {\normalfont\small\bfseries}}
\makeatother

% For pmatrix  equal spacing:
\renewcommand*{\arraystretch}{1.1}
\renewcommand*{\arraycolsep}{1.2pt}

\setcounter{secnumdepth}{0}
\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0.5ex}
\setlength{\parskip}{0.02em}%Length of space between paragraphs

\setlength{\columnsep}{6mm}

\setlength{\itemsep}{0.2em}

% -----------------------------------------------------------------------

\title{My Title}

\begin{document}

\raggedright
\footnotesize

\begin{center}
     \Large{\textbf{My Title}}\\\small{by Sebastian Krantz} %\vspace{mm}
\end{center} 
\begin{multicols*}{4}
\setlength{\premulticols}{1pt}
\setlength{\postmulticols}{1pt}
\setlength{\multicolsep}{1pt}
\setlength{\columnsep}{2pt}

\section{\textcolor{Gray}{Basics}}
\hline

collapse is a powerful (C/C++ based) package supporting advanced (grouped, weighted, time series, panel data and recursive) operations in R. \newline

It also offers a class-agnostic approach to data manipulation - handling matrix and data frame based objects in a uniform \& non-destructive way. \newline

It is particuarly well integrated with data.table, dplyr ((grouped) tibbles) and plm classes for panel data, and can be programmed using pipes \%>\% (magrittr) or standard evaluation.

\section{Fast Statistical Functions}
bla bla

\end{multicols*}
\end{document}